---
title             : "A Comparative Analysis on Native and Non-Native Korean Speakers' Vowel Productions"
shorttitle        : "Korean Vowels"

author: 
  - name          : "Jungah-Lee"
    affiliation   : "1"
    corresponding : yes    # Define only one corresponding author
    address       : "1248 University of Oregon Eugene, Oregon 97403-1248"
    email         : "jlee27@uoregon.com"
  - name          : "Heidi Shi"
    affiliation   : "2"
  - name          : "Jeongim Jin"
    affiliation   : "3"

affiliation:
  - id            : "1"
    institution   : "University of Oregon"
  - id            : "2"
    institution   : "University of Oregon"
  - id            : "3"
    institution   : "University of Oregon"

authornote: |
  Jung-ah Lee joined the Department of East Asian Langauges and Literatures at the Unviersity of Oregon at 2017.


abstract: |

 This paper investigates Korean vowel spaces of native Korean speakers (NS) and non-native Korean speakers (NNS, L1: English). The vowel production of NS and NNS were compared. All the speakers recorded eight Korean cardinal vowels, which are [i], [e], [ae], [W], [A], [o], [u], and [a] three times. A preceding consonant was lenis fricative [s] sound of Korean. The participants’ production was normalized by using Lobanov normalization function in the R program. Results show that NNS cannot differentiate the Korean mid-vowels and back vowels such as [A], [o], and [u]. This research can be helpful to investigate that L1 English can affect the L2 Korean learners’ vowel pronunciation.   
 
  <!-- https://tinyurl.com/ybremelq -->
  
keywords          : "Community of Practice, Language Socialization"
wordcount         : "X"

bibliography      : ["Group 5 final-references.bib"]

floatsintext      : no
figurelist        : no
tablelist         : no
footnotelist      : no
linenumbers       : yes
mask              : no
draft             : no

documentclass     : "apa6"
classoption       : "man, fleqn, noextraspace"
output            : 
                   papaja::apa6_pdf:
                     toc: true
                     toc_depth: 4
                     latex_engine: xelatex
                    
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(echo = FALSE, 
                      message = FALSE,
                      warning = FALSE,
                      fig.width = 6.5,
                      fig.height = 8)

library(vowels)
library(here)
library(rio)
library(languageR)
library(tidyverse)
library(janitor)
library("papaja")


theme_set(theme_minimal(base_size = 8))

```

```{r analysis-preferences}
# Seed for random number generation
set.seed(42)
knitr::opts_chunk$set(cache.extra = knitr::rand_seed)
```


\newpage

# **1.Introduction**
In this paper, our group investigated Korean vowel productions of Korean native speakers (NS) and non-native Korean speakers (NNS) by analyzing acoustic features. We aimed to examine native speaker pattern of vowel formants in Korean, and to explore second language learners’ patterns in comparison to those of NNSs. Our results were analyzed by using normalization and statistic functions in the R program. The results indicate that L1 (English) of the NNSs might affect their L2 Korean vowel production. It is important to know what sounds students may have trouble with and what causes the difficulty in L2 Korean acquisition (Idemaru, 2018). Thus, this study is useful in terms of as well as suggesting crucial research in the second language education field. 

# **2. Methodology**
<<<<<<< HEAD
To compare the convergence and divergence of L1 Korean and L2 Korean Vowel Production, we used a one-way between-subjects design, and analyzed the collected data using two Hz-basd measurements and a phonetic measurement tool, Praat. 
=======
To compare the convergence and divergence of L1 Korean and L2 Korean Vowel Production, we used a one-way between-subjects design, and analyzed the collected data using a one-way ANOVA.
>>>>>>> master

## *2.1 Participants*
Using convenince sampling, we sampled six adult residents living in Eugene, Oregon; three of them were Native Koreans, and the rest were non-Native sperking Koreans. Two Native Koreans were doctoral students majoring in East Asian Lanuages and Literatures, and the other one was a Korean instructor at the University of Oregon. Non-native speakers were recruited from one of the Korean classes at the Univeresity of Oregon. Korean participants were an average of 28 in age, while the average age of the non-Korean subjects were 19. 

## *2.2 Speech Materials*
We prepared seven sentences consisting of eight Korean vowels; i,e,ae,w,^,a,u. The sentences to be tested will be added at the end of the orignial paper as one of the appendice. 

## *2.3 Procedure*
The Korean vowles of the participants were audie-recorded. We met in the quite room individally and we had them repeat the eight sentences three times using a Praat program, which is widely used in phonetic measurements. By measuring F1 and F2 in the vowel mid-point, we intended to find vowel formant patterns. 

# **3. Restuls and Discussion**

# **4. Conclusion**

\newpage

# **References**
```{r create_r-references}
r_refs(file = "Group 5 final-references.bib")
```

\begingroup
\setlength{\parindent}{-0.5in}
\setlength{\leftskip}{0.5in}

<div id = "refs"></div>
\endgroup
