---
title             : "A Comparative Analysis on Native and Non-Native Korean Speakers' Vowel Productions"
shorttitle        : "Korean Vowels"

author: 
  - name          : "Jungah-Lee"
    affiliation   : "1"
    corresponding : yes    # Define only one corresponding author
    address       : "1248 University of Oregon Eugene, Oregon 97403-1248"
    email         : "jlee27@uoregon.com"
  - name          : "Heidi Shi"
    affiliation   : "2"
  - name          : "Jeongim Jin"
    affiliation   : "3"

affiliation:
  - id            : "1"
    institution   : "University of Oregon"
  - id            : "2"
    institution   : "University of Oregon"
  - id            : "3"
    institution   : "University of Oregon"

authornote: |
  Jung-ah Lee joined the Department of East Asian Langauges and Literatures at the Unviersity of Oregon at 2017.


abstract: |

  
  <!-- https://tinyurl.com/ybremelq -->
  
keywords          : "Community of Practice, Language Socialization"
wordcount         : "X"

bibliography      : ["Group 5 final-references.bib"]

floatsintext      : no
figurelist        : no
tablelist         : no
footnotelist      : no
linenumbers       : yes
mask              : no
draft             : no

documentclass     : "apa6"
classoption       : "man, fleqn, noextraspace"
output            : 
                   papaja::apa6_pdf:
                     toc: true
                     toc_depth: 4
                    
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(echo = FALSE, 
                      message = FALSE,
                      warning = FALSE,
                      fig.width = 6.5,
                      fig.height = 8)

install.packages("rio")
install.packages("here")
install.packages("vowels")
install.packages("languageR")

library(vowels)
library(here)
library(rio)
library(languageR)
library(tidyverse)
library(janitor)
library("papaja")


theme_set(theme_minimal(base_size = 8))

```

```{r analysis-preferences}
# Seed for random number generation
set.seed(42)
knitr::opts_chunk$set(cache.extra = knitr::rand_seed)
```


\newpage

# **1.Introduction**
We used @pan2011 XXXX

# **2. Methodology**
We adopted the method introduced by @pan2011, the author of *Phonetics and phonology of Nantong Chinese*We used methods introduced by @ahmed2001. Besides we also adopted the diligent and collaborative approach to Rstudio (see @kadar2011).

## *2.1 Participants*
The authors gathered teenagers who attend different schools on the Pacific Northwest, *Eugene,Oregon*.

### Korean Native Speakers ###
n=3

### Korean Non-native Speakers ###
n=3

## *2.2 Speech Materials*

## *2.3 Procedure*

# **3. Restuls and Discussion**

# **4. Conclusion**

\newpage

# **References**
```{r create_r-references}
r_refs(file = "Group 5 final-references.bib")
```

\begingroup
\setlength{\parindent}{-0.5in}
\setlength{\leftskip}{0.5in}

<div id = "refs"></div>
\endgroup
